---
title: "Report"
author: "Oscar Leal"
date: "3/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(Hmisc)
library(ggplot2)
library(dplyr)
library(tidyr)
library(janitor)

library(DataExplorer)
```


## Introduction of *http://archive.ics.uci.edu/ml/machine-learning-databases/00502/* dataset.

Transactions occurring for a UK-based and registered, non-store online retail between 01/12/2009 and 09/12/2010.

Changing format for a more lightweight solution.

```{r}
#data <- readxl::read_xlsx("../online_retail_II.xlsx")

#saveRDS(data, file = "online_retail_II.rds")

```

Reading data.

```{r, include = FALSE}
data <- readRDS("online_retail_II.rds")
```

```{r}
colnames(data)
```

InvoiceNo: Invoice number. Nominal. A 6-digit integral number uniquely assigned to each transaction. If this code starts with the letter 'c', it indicates a cancellation.

StockCode: Product (item) code. Nominal. A 5-digit integral number uniquely assigned to each distinct product.

Description: Product (item) name. Nominal.

Quantity: The quantities of each product (item) per transaction. Numeric.

InvoiceDate: Invoice date and time. Numeric. The day and time when a transaction was generated.

UnitPrice: Unit price. Numeric. Product price per unit in sterling (£).

CustomerID: Customer number. Nominal. A 5-digit integral number uniquely assigned to each customer.

Country: Country name. Nominal. The name of the country where a customer resides.

```{r}
# some basic cleaning and factoring these columns for easier visualization of ratios
data <- data %>% clean_names()
data[, 'description'] <- as.factor(data[, 'description'])
data[, 'country'] <- as.factor(data[, 'country'])
data[, 'customer_id'] <- as.factor(data[, 'customer_id'])
data[, 'stock_code'] <- as.factor(data[, 'stock_code'])



summary(data)
tail(data)
str(data)

```

## Comparing daily turnover of online retailer platform in 2009 and 2010 December.


