---
title: "Report"
author: "Oscar Leal"
date: "3/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(Hmisc)
library(ggplot2)
library(dplyr)
library(tidyr)
library(janitor)
library(plotly)
library(TSstudio)
library(data.table)
library(lubridate)


library(DataExplorer)
```


## Introduction of *http://archive.ics.uci.edu/ml/machine-learning-databases/00502/* dataset.

Transactions occurring for a UK-based and registered, non-store online retail between 01/12/2009 and 09/12/2010.

Changing format for a more lightweight solution.

```{r}
#data <- readxl::read_xlsx("../online_retail_II.xlsx")

#saveRDS(data, file = "online_retail_II.rds")

```

Reading data.

```{r, include = FALSE}
data <- readRDS("online_retail_II.rds")
```

```{r}
colnames(data)
```

InvoiceNo: Invoice number. Nominal. A 6-digit integral number uniquely assigned to each transaction. If this code starts with the letter 'c', it indicates a cancellation.

StockCode: Product (item) code. Nominal. A 5-digit integral number uniquely assigned to each distinct product.

Description: Product (item) name. Nominal.

Quantity: The quantities of each product (item) per transaction. Numeric.

InvoiceDate: Invoice date and time. Numeric. The day and time when a transaction was generated.

UnitPrice: Unit price. Numeric. Product price per unit in sterling (£).

CustomerID: Customer number. Nominal. A 5-digit integral number uniquely assigned to each customer.

Country: Country name. Nominal. The name of the country where a customer resides.

```{r}
# some basic cleaning and factoring these columns for easier visualization of ratios
data <- data %>% clean_names()
data$description <- as.factor(data$description)
data$country <- as.factor(data$country)
data$customer_id <- as.factor(data$customer_id)
data$stock_code <- as.factor(data$stock_code)


summary(data)
tail(data)
str(data)

#data explorer report
#DataExplorer::create_report(data)

```

```{r}
#here we notice that quantity can be negative. this indicates it should be a good idea to create a total price value that multiplies qty * price
data %>% filter(startsWith(invoice,"C")) %>% filter(stock_code == "22087")


data <- data %>% mutate(total_price = quantity * price)
```

## Comparing daily turnover of online retailer platform in 2009 and 2010 December.


```{r, echo = FALSE}
# filters for only december
decembercomparison2009 <- data %>% filter(month(invoice_date) == 12 & (year(invoice_date) %in% c(2009))) %>% mutate(day = lubridate::day(invoice_date), year = year(invoice_date)) %>% group_by(day, year) %>% summarise(daily_sales=sum(total_price)) %>% ungroup()

decembercomparison2009 <- add_row(decembercomparison2009, day = 24:31, year = 2009, daily_sales = 0)

decembercomparison2010 <- data %>% filter(month(invoice_date) == 12 & (year(invoice_date) %in% c(2010))) %>% mutate(day = lubridate::day(invoice_date), year = year(invoice_date)) %>% group_by(day, year) %>% summarise(daily_sales=sum(total_price)) %>% ungroup()

decembercomparison2010 <- add_row(decembercomparison2010, day = 10:31, year = 2010, daily_sales = 0)

decembercomparison2009 <- decembercomparison2009 %>% mutate( date = as.Date(ymd(paste0(year, "-12-", day)))) %>% select(daily_sales, date)
decembercomparison2010 <- decembercomparison2010 %>% mutate( date = as.Date(ymd(paste0(year, "-12-", day)))) %>% select(daily_sales, date)
```


```{r}
ts_plot(decembercomparison2009)
```

```{r}
ts_plot(decembercomparison2010)
```


## Visualizing distribution of daily purchases

```{r}


```






```{r}
theme_set(theme_bw())

# Draw plot
data %>%
  ggplot(aes(class, cty)) %>% 
  geom_boxplot(varwidth=T, fill="plum") + 
    labs(title="Box plot", 
         subtitle="City Mileage grouped by Class of vehicle",
         caption="Source: mpg",
         x="Class of Vehicle",
         y="City Mileage")
```
